---
title: "Assignment_week9"
author: "Umais Siddiqui"
date: "October 29, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

http://rpubs.com/umais/data607_week9_assignment

https://github.com/umais/DATA-607/tree/master/Week9_Assignment

# Web API's

##Approach

Using the jsonLite library we will be reading the JSON document from the New your times API. Once document is read we will transform the json in to a R data frame object

##Reading the data

After retrieving the API Key Making a call to the Ny Times Article search API to get the listing of the articles
```{r}
library(jsonlite)
nytimes_articles <- fromJSON("http://developer.nytimes.com/proxy/https/api.nytimes.com/svc/search/v2/articlesearch.json?api-key=d0aaa22b5c5e46dea282c685d7ecfb07")

```




##Transforming it in to a data frame.

In this step we successfully created a data frame from the JSON returned from the API call
```{r}
articleDf=data.frame(nytimes_articles$response$docs$pub_date)
colnames(articleDf)<-c('publishdate')
articleDf$type=nytimes_articles$response$docs$document_type
articleDf$headline=nytimes_articles$response$docs$headline$main
articleDf$source=nytimes_articles$response$docs$source
articleDf$section_name=nytimes_articles$response$docs$section_name
articleDf$web_url=nytimes_articles$response$docs$web_url

```


##Display the data in the data frame.

```{r}
articleDf

```



