---
title: "Assignment 2"
author: "Umais Siddiqui"
date: "September 4, 2017"
output:
  pdf_document: default
  html_document: default
---

Git hub : https://github.com/umais/DATA-607/tree/master/Assignment2"

Rpub : http://rpubs.com/umais/305818"

#Movies Database

This R Markdown is written to work with the MYSQL Movies database. The SQL file can be viewed from the githib link below.  

Th database that is created has three tables . RecentMovies, Person and PersonMovieRating. RecentMovies will store the movie information for the recent movies. The PErson table stores all the people that were surveyed to rate the movie. The PersonMovieRating stores the ratings that were collected by the people in the person table. We will be joining these tables and gathering this information in one select query and then loading it in to a R data frame.




```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```
##Step 1 Load the R MYSQL Library
```{r}

library(RMySQL)
```

##Step 2 Connect to the MySQL Movies database 

     In this step we will be connecting to the Movies database and retrieving the information that
     was collected in the survey and load that in the R Data Frame
```{r }
mydb = dbConnect(MySQL(), user='root', password='Welcome@1', dbname='movies', host='localhost')

rs = dbSendQuery(mydb, "SELECT m.MovieName,p.FirstName,p.LastName,r.Rating 
FROM PersonMovieRating r INNER JOIN RecentMovie m 
on r.MovieId=m.MovieId
INNER JOIN Person p
ON r.personID=p.personid;")

df=fetch(rs, n=-1)
```

##Step 3 Show the first few rows using the head function and show the summary of the ratings.
```{r}
head(df)
summary(df$Rating)
```