---
title: "Week 1 Assignment"
author: "Umais Siddiqui"
date: "September 2, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mushrooms Dataset 

##Step 1 Import Libraries

```{r }
library(plyr)
library(RCurl)
```
##Step 2 Reading from CSV file located in Github repo
In this step we will be reading the data from csv file in to a data frame and then renaming the columns in the data frame to more meaningful names.

```{r df}
url<-getURL("https://raw.githubusercontent.com/umais/DATA-607/master/Assignment1/agaricus-lepiota.data")

df <- rename(data.frame(read.csv(text=url,header=FALSE)),
                        c(
                          "V1"="classes",
                          "V2"="cap_shape",
                          "V3"="cap_surface",
                          "V4"="cap_color",
                          "V5"="Bruises"
                          ))[,c(1:5)]

head(df)
```

## Step 3 Replace Values in the columns and Display the data frame

In this step we will be replace the values of the data to have more menaingful values

```{r}
revalue(df$classes, c("e" = "edible")) -> df$classes
revalue(df$classes, c("p" = "poisonous")) -> df$classes
revalue(df$cap_shape, c("b" = "bell")) -> df$cap_shape
revalue(df$cap_shape, c("c" = "conical")) -> df$cap_shape
revalue(df$cap_shape, c("x" = "convex")) -> df$cap_shape
revalue(df$cap_shape, c("f" = "flat")) -> df$cap_shape
revalue(df$cap_shape, c("k" = "knobbed")) -> df$cap_shape
revalue(df$cap_shape, c("s" = "sunken")) -> df$cap_shape
revalue(df$cap_surface, c("f" = "fibrous")) -> df$cap_surface
revalue(df$cap_surface, c("g" = "grooves")) -> df$cap_surface
revalue(df$cap_surface, c("y" = "scaly")) -> df$cap_surface
revalue(df$cap_surface, c("s" = "smooth")) -> df$cap_surface
revalue(df$cap_color, c("n" = "Brown")) -> df$cap_color
revalue(df$cap_color, c("b" = "Buff")) -> df$cap_color
revalue(df$cap_color, c("c" = "Cinnamon")) -> df$cap_color
revalue(df$cap_color, c("g" = "Gray")) -> df$cap_color
revalue(df$cap_color, c("r" = "Green")) -> df$cap_color
revalue(df$cap_color, c("p" = "Pink")) -> df$cap_color
revalue(df$cap_color, c("u" = "Purple")) -> df$cap_color
revalue(df$cap_color, c("e" = "Red")) -> df$cap_color
revalue(df$cap_color, c("w" = "White")) -> df$cap_color
revalue(df$cap_color, c("y" = "Yellow")) -> df$cap_color
revalue(df$Bruises, c("t" = "Yes")) -> df$Bruises
revalue(df$Bruises, c("f" = "No")) -> df$Bruises
df
```
