---
title: "Project2 Data 607"
author: "Umais Siddiqui"
date: "October 7, 2017"
output:
  html_document: default
  pdf_document: default
---




```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(RMySQL)
library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

###Rpubs Link: http://rpubs.com/umais/data607_Project2 

###Github Link: https://github.com/umais/DATA-607/tree/master/Project2_Data607

#Down Stream Analysis using tidyR and Dplyer

I will be using three data sets from a past projects that I worked on. One of the data sets is from the company I work for at the moment.


##School Data Set
1) Data set from a schools project I did for couple of schools back home. The data shows the number of students that passed and failed from 9th and  10th grade . I have deindentified the data by using fake school names
```{r }
mydb = dbConnect(MySQL(), user='root', password='Welcome@1', dbname='project2', host='localhost')

rs = dbSendQuery(mydb, "SELECT * FROM StudentRecords;")

df=fetch(rs, n=-1)

head(df)
```

###Summarise the Data in to total Passed and Failed from each school
```{r}
#Gather Function from tidyR
TotalStudents<- gather(df,Status,NumberOfStudents,NinthGradePass:TenthGradeFail)

head(TotalStudents)

failed=TotalStudents %>% dplyr::filter(Status %like% "Fail")%>%group_by(SchoolName)%>%summarise(total=sum(NumberOfStudents))

failed

passed=TotalStudents %>% dplyr::filter(Status %like% "Pass")%>%group_by(SchoolName)%>%summarise(total=sum(NumberOfStudents))

passed


       
```

##Second Data Set for In Network and Out of Network Providers 

```{r}

rs = dbSendQuery(mydb, "SELECT * FROM ProviderNetwork ;")

df=fetch(rs, n=-1)

head(df)

TotalProviders<- gather(df,NetworkStatus,NumberOfProviders,TotalInNetworkChicago:TotalOutOfNetwork_NewYork)

TotalProviders



```

###Check Which Health Plan has More In Network Providers from the two cities.

```{r}
InNetwork=TotalProviders %>% dplyr::filter(NetworkStatus %like% "InNetwork")%>%group_by(PayerName)%>%summarise(total=sum(NumberOfProviders))

InNetwork

```

As we can See Payer 1 has more in network providers than payer 2.

##Using the Third data set for the MemberRisk Data

```{r}


rs = dbSendQuery(mydb, "SELECT * FROM  MemberRisk ;")

df=fetch(rs, n=-1)

head(df)

TotalMembers<- gather(df,RiskStatus,NumberOfMembers,HighRiskFlorida:LowRiskNewYork)

TotalMembers

HighRisk=TotalMembers %>% dplyr::filter(RiskStatus %like% "High")%>%group_by(OrganizationName)%>%summarise(total=sum(NumberOfMembers))

HighRisk
```

In the last Data Set we can see that AlignCare has more High Risk Members than Demo Data Large


