---
title: "Week12_Assignment"
author: "Umais Siddiqui"
date: "November 16, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mongolite)
library(RMySQL)
```

Rpubs: http://rpubs.com/umais/week12_assignment

Github: 

## Assignment Week 12

In this assignment I will be taking data from a relational database and migrating it to MongoDB database that is set up on MongoLab.
The data set that I have chosen is from Project 3 . I am going to retrieve the Job Title, Job Location and Skill Name from the Normalized tables DataScienceJob and DataScienceSkills and store them in a mongo database collection called jobskills.

### Reading the Data from MySQL database and migrating it to MongoDB No SQL Storage

In this step I will be using the RMySQL library to connect to MySQL data store and retrieve the Skills data. I will then connect to the MongoDB database called openletter on MongoLab and Insert the Skill data reteived from MySQL in the collection called jobskills.

```{r}
my_collection = mongo(collection = "jobskills", url = "mongodb://ruser:Tw1stone23@ds031618.mlab.com:31618/openletter")
mydb = dbConnect(MySQL(), user='root', password='Welcome@1', dbname='project3', host='localhost')

results = dbSendQuery(mydb, "SELECT j.JobTitle,j.JobLocation,s.SkillName FROM DataScienceJobs j INNER JOIN DataScienceSkills s ON j.JobId=s.JobId  ;")

jobSkills=fetch(results, n=-1)
my_collection$drop()
my_collection$insert(jobSkills)
head(jobSkills)

```

### Reading data from the MongoDB collection to ensure it got Saved correctly.

```{r}
alldata <- my_collection$find('{}')
head(alldata)

```

### Pros and Cons of Storing data in Relational DB vs NoSQL

1) The advanatage of storing data in a relational database is that data can be stored in a relational manner that is good for data integrity and elimnates alot of anomolies.
2) The advantage of storing data in a NoSQL store is that it is easier to retrieve and the data access is faster as the data is stored in a manner of document.

