---
title: "Week 5 Assignment"
author: "Umais Siddiqui"
date: "September 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(RMySQL)
knitr::opts_chunk$set(echo = TRUE)
```

## Tidying and Transforming Data


###Assignment Overview
In this assignment we will be using the TidyR and Dplyer packages to perform some analysis on the delayed arrival for two Airlines.

###Data Set

The data set that we will be using will be from MySQL data source. First step would be to use the RMySQL package to connect to the MySQL database and retrieve the data.



```{r }
mydb = dbConnect(MySQL(), user='root', password='Welcome@1', dbname='flights', host='localhost')

rs = dbSendQuery(mydb, "SELECT * FROM AirlineArrival;")

df=fetch(rs, n=-1)

head(df)
```

###Using tidyR to transform the data

As we can see from the results of the data that the data set is in a wide format. What we would like to do in our first step is transform the data in to a long format so that we have the following columns Airline,ArrivalStatus,Cities, NumberOfFlights. We can achive this using the tidyR function called gather.

```{r}
#Gather Function from tidyR
airlines2<- gather(df,Cities,NumberOfFlights,LosAngeles:Seattle)

head(airlines2)
```

###Using Dplyer to filter results

```{r}
#Usinf the FIlter Function from DPlyer
delayedFlights=filter(airlines2,ArrivalStatus=="Delayed")
delayedFlights
```


###Using DPlyer Pipeline and Mutate function to look at percentages of flights delayed to 